---
title: "Exercise 1.5"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require('knitr')
opts_knit$set(root.dir = "~/Documents/STEMinist_R/day_1/Day_1_Exercises/")
```
### Data clean up 
#### useful commands: `duplicated()`, `is.na()`, `complete.cases()`, `!duplicated()`

##### load the sleep data as in exercise 1.4
##### this data describes properties of sleep for different mammals
##### column names are:
- BodyWgt   body weight (kg)
- BrainWgt  brain weight (g)
- NonD      slow wave (non-dreaming) sleep (hours/day)
- Dream     paradoxical (dreaming) sleep (hours/day)
- Sleep     total sleep (hours/day)
- Span      maximum life span (years)
- Gest      gestation time (days)
- Pred      Predation index (1-5) 1 = minimum (least likely to be preyed upon), 5 = max
- Exp       sleep exposure index (1-5) 1 = least exposed (sleeps in well-protected den), 5 = most exposed
- Danger    overall danger index (1-5) 1 = least danger (from other animals), 5 = most danger

```{r}
sleep = read.csv('sleep.csv')
```


#* 01. How many duplicated rows are in the dataset?

```{r}
sum(duplicated(sleep))
```


# 02. Remove the duplicated rows from the datset

```{r}
dim(sleep)
sleep = sleep[!duplicated(sleep),]
dim(sleep)  ## note that there are six fewer rows now
```


#* 03. Subset the dataset to only include records with data on Gestation period

```{r}
sleep[!is.na(sleep[,'Gest']),]  ## so we are only taking rows where Gest is not (!) NA
```


# 04. Subset the dataset to only include records with data on Dreaming and non dreaming sleep

```{r}
sleep[!is.na(sleep[,'Dream']) & !is.na(sleep[,'NonD']),]
```


#* 05. Subset the dataset to only include complete records

```{r}
sleep[complete.cases(sleep),]  ## remember that complete.cases() returns indexes of rows
```

# 06. How heavy is the heaviest mammal in the dataset?

```{r}
max(sleep[,'BodyWgt'])
```


#* 07. What is the brain weight of the heaviest individual?
```{r}
sleep[,'BrainWgt'][which.max(sleep[,'BodyWgt'])]
```


# 08. What are the weights of the animals that share the most common combination of Pred/Exp/Danger? 
#      What is the mean weight of those animals? (This is hard. Don't worry if you don't get it.)

```{r}
table(sleep[,c('Pred', 'Exp', 'Danger')]) ## this will give us a three way table
max(table(sleep[,c('Pred', 'Exp', 'Danger')]))  ## this will tell us which is the max
## look in the table above to see this value of 10 is observed when Danger==1, Pred==1, and Exp==1
sleep[,'BodyWgt'][sleep[,'Pred']==1 &   ## we can spread this command across lines when it gets unwieldy
                    sleep[,'Danger']==1 &
                    sleep[,'Exp']==1]

mean(sleep[,'BodyWgt'][sleep[,'Pred']==1 &  
                    sleep[,'Danger']==1 &
                    sleep[,'Exp']==1]
)
```


This data came from a paper by [Allison and Cicchetti, 1976, \textit{Science}](http://seek.ecoinformatics.org/attach%3Fpage=Ecoinformatics2003%252Fallison+and+cicchetti+article.pdf)

For fun, these least preyed upon, least in danger, and least exposed when sleeping animals are:

- Arctic fox
- Big brown bat
- Chimpanzee
- Eastern American mole
- Giant armadillo
- Gray wolf
- Jaguar
- Little brown bat
- Man
- Red fox