---
title: "Exercise 1.1"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Interacting with R 
#### useful functions: `c()`, `ls()`, `rm()`, `objects()`


##### * 01. Open Rstudio and perform an arithmetic calculation in the command line.
```{r question1}
1 + 1
2 * 8
```

##### * 02. Create a numeric vector in the command line containing:

- the numbers 2, 9, 3, 8, and 3 and assign this vector to a global variable x. 

```{r}
x = c(2,9,3,8,3)
x
```


- Perform arithmetic with x. 

```{r}
x + 2000
```


- Convince yourself R works as a calculator, and knows order of operations. 

```{r}
1 + 3 * 10^2
```


- Multiply x by 10, and save the result as a new object named y

```{r}
y = x * 10
```

        
- Calculate the difference in the sum of the x vector and the sum of the y vector

```{r}
sum(x) - sum(y)
```

##### 03. Call the help files for the functions ls() and rm()
```{r}
?ls
help(ls)
```

- What are the arguments for the ls() function?

The arguments for the ls() function are: name, pos, envir, all.names, pattern, and sorted.

- What does the 'sorted' argument do? 
       
The sorted argument takes a logical value, which denotes whether the vector should be sorted alphabetically or not. 
By default, ls() uses sorted=TRUE.

##### 04. List all objects in the global environment. 

```{r}
ls()
```

- Remove x and y from the global environment.
```{r}
rm(x)
rm(y)
```

