---
title: "Exercise 1.2"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Characterize a data frame

useful commends: `[row , col]`, `str()`, `dim()`, `nrow()`, `unique()`, `length()`, `rownames()`, `summary()`, `min()`, `max()`, `mean()`, `range()`, `levels()`, `factor()`, `as.factor()`, `class()`, `ncol()`, `nrow()`, `table()`, `sum()`, `quantile()`, `var()`

## We'll use the built-in 'iris' dataset
```{r}
data(iris) # this loads the 'iris' dataset 
```

# load help file for the 'iris' dataset

```{r}
?iris
help(iris)
```

# this calls the object iris
```{r}
head(iris) ## let's not print the entire thing to our screens
```

If you're not a botanist, remind yourself what an iris looks like http://suruchifialoke.com/img/ML/iris.png

#* 01. How many rows are in the dataset?

```{r}
nrow(iris)
```


#  02. What are three distinct ways to figure this out? 
```{r}
nrow(iris)
dim(iris)[1]
length(rownames(iris))
str(iris)
```


#* 03. How many species of flowers are in the dataset?

```{r}
length(unique(iris[,'Species'])) ## this counts how long the unique() vector is
length(table(iris[,'Species'])) ## this asks how many entries are in the table
```

# 04. What class is iris?
```{r}
class(iris)
```

#* 05. How many columns does this data frame have? What are their names?

```{r}
ncol(iris)
colnames(iris)
```

#* 06. What class did R assign to each column?

```{r}
class(iris[,1])
class(iris[,2])
class(iris[,3])
class(iris[,4])
class(iris[,5])
```
Or, we can use `lapply()`, which we'll learn about later
```{r}
## ssh, we're not supposed to learn this until Day 3
lapply(iris, class)
```


# 07. Assign the first flower's petal's width and length to new objects called setosa1.petalwidth and setosa1.petallength

```{r}
setosa1.petalwidth = iris[1,'Petal.Width']
setosa1.petallength = iris[1,'Petal.Length']
```


#* 08. Calculate the approximate area of the petal of the first flower, setosa1 (assume petal area can be approximated by a rectangle).

```{r}
setosa1.petalwidth * setosa1.petallength
```


# 09. Calculate the petal area of each flower in the iris dataset
#        and assign this to a new column named PetalArea.

```{r}
iris[,'PetalArea'] = iris[,'Petal.Width'] * iris[,'Petal.Length']
head(iris)
```


# 10. What is the maximum sepal length of the irises?

```{r}
max(iris[,'Sepal.Length'])
```


#* 11. What is the average sepal length among flowers in the datset?

```{r}
mean(iris[,'Sepal.Length'])
```


# 12. How about the minimum and median sepal length? 

```{r}
min(iris[,'Sepal.Length'])
median(iris[,'Sepal.Length'])
```


