---
title: "Exercise 1.4"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require('knitr')
opts_knit$set(root.dir = "~/Documents/STEMinist_R/day_1/Day_1_Exercises/")
```
### Getting data into R
#### useful commands `setwd()`, `read.csv()`, `list.files()`, `getwd()`

##### * 01. Enter getwd() to determine your current working directory. 
  
```{r}
getwd()
```

    
# *02. Set working directory to the class folder you have from Google Drive 

```{r, eval=FALSE, include=TRUE}
## this depends on where you've saved things, and will be different for everybody!
setwd('~/Documents/STEMinist_R/day_1/Day_1_Exercises/')  
```


#* 03. List the files present in this directory.

```{r}
list.files()
```


#* 04. Read "sleep.csv" into R, store as an object sleep

```{r}
sleep = read.csv('sleep.csv')
```

  
# 05. How many columns are in sleep?

```{r}
ncol(sleep)
```


# 06. Remove sleep from your R working session

```{r}
rm(sleep)
```

  
# 07. Move up one working directory to the parent folder,
#       list objects present, and then move back down to the course folder.
       
    
```{r, eval=FALSE, include=TRUE}
setwd('..')  # in unix, .. means up one directory, . means the current directory
list.files()
## note that unless you've named your directories just like mine, this may be different!
setwd('Day_1_Exercises')  
```
