---
title: "Exercise 1.3"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Indexing by logical statements
#### useful commands:  `==`, `!=`, `>`, `<`, `&`, `|`, `sum()`, `which()`, `table()`, `!`

##### * 1. Create your own logical vector with three TRUEs and three FALSEs

```{r}
a = c(TRUE, TRUE, FALSE, FALSE, TRUE, FALSE)
a ## let's print to screen and make sure it is stored in this variable
```


##### * 2. Produce a vector of the index number of the true’s
```{r}
which(a)  ## which gives you the index of TRUE values automatically
which(a == TRUE)   ## but sometimes it's reassuring to state exactly what you're doing
```


#####   3. Produce a second vector which indexes the numbers of the falses
```{r}
which(!a)
which(a == FALSE)
```


## Go back to the iris dataset

```{r}
data(iris)  ## this reloads the data set in case you've closed R since using iris
```


#####  4. How many irises have sepals less than 5.5 cm long?
```{r}
sum(iris[,'Sepal.Length']<5.5)  ## remember TRUE's are 1 and FALSE's are 0
length(which(iris[,'Sepal.Length']<5.5))  ## here, which() will only return the index of TRUE values, so we're counting how many there are
```


##### * 5. Which iris individual has the largest petal length? What is the width of it's petal? 

```{r}

max(iris[,'Petal.Length'])  ## this gives us the length of the longest petal
which(iris[,'Petal.Length'] == max(iris[,'Petal.Length']))  ## this gives us the index of the individual with the longest petal
iris[,'Petal.Width'][which(iris[,'Petal.Length'] == max(iris[,'Petal.Length']))] ## now we're subsetting the Petal.Width column by the index of the individual with the longest petal
## another way to do this would be to use the index of the individual with the longest petal to pick rows, and the Petal.Width name to pick columns and subset the entire data frame
iris[which(iris[,'Petal.Length'] == max(iris[,'Petal.Length'])) , 'Petal.Width']
```


#####   6.  How many of the irises are in this dataset belong to the species versicolor?

```{r}
sum(iris[,'Species']=='versicolor')
table(iris[,'Species']) ## this could get us close
```


##### * 7. How many irises have petals longer than 6cm?
```{r}
sum(iris[,'Petal.Length'] > 6)
```


#####   8. Create a vector of species name for each iris with sepals longer than 6cm.

```{r}
iris[,'Species'][iris[,'Sepal.Length']>6]
iris[iris[,'Sepal.Length']>6, 'Species'] ## alternatively, we can put the logical vector in the row part, and Species in the column part, to get a vector back
```


##### * 9. How many irises have sepals shorter than 5cm, but wider than 3cm?
```{r}
sum( iris[,'Sepal.Length'] < 5 & iris[,'Sepal.Width'] > 3 )
```


##### 10. How many irises have petals narrower than 0.2cm or shorter than 1.5cm?

```{r}
sum( iris[,'Petal.Width'] < 0.2 | iris[,'Petal.Length'] < 1.5 )
```

##### 11. What is the average width of setosa iris sepals that are longer than 5cm?

```{r}
mean( iris[,'Sepal.Width'][iris[,'Sepal.Length'] > 5]) ## convince yourself the second part is a logical vector that subsets iris[,'Sepal.Width']
mean( iris[iris[,'Sepal.Length'] > 5, 'Sepal.Width']) ## again, we can alternatively subset using logical vectors in the row position
```


##### 12. What is the difference between the longest and shortest petal lengths of the species virginica?
```{r}
max(iris[,'Petal.Length'][iris[,'Species']=='virginica']) - min(iris[,'Petal.Length'][iris[,'Species']=='virginica'])
```


##### 13. What proportion of flowers in the dataset have petals wider than 1cm?

```{r}
sum(iris[,'Petal.Width'] > 1 ) / nrow(iris) ## here, we're counting up how many are wider than 1 cm, and dividing by the total number of flowers to get a proportion
```


##### 14. Create a new column within your dataframe, called sepalCategory, and set all values equal to 'long'
```{r}
iris[,'sepalCategory'] = 'long'  ## this sets ever entry in the column equal to 'long'
```


- Subset short values of this column, and set their values to 'short'
 (Short sepals are those less than 5.5 cm)
       
```{r}
iris[,'sepalCategory'][iris[,'Sepal.Length']<5.5] = 'short'  ## this sets only those entries that match our condition to 'short'

```

- How many plants with short sepals are there? How many long?


```{r}
table(iris[,'sepalCategory'])
```

