---
title: "Chick Weight Challenge"
output:
  pdf_document: default
  word_document: default
---

In this challenge exercise, we're going to explore a dataset from an experiment on the effect of diet on early growth of chicks. At the end, we will use two useful packages: R Color Brewer which creates nice color palettes and R Markdown to turn our analysis into a final report.

### Download and read the comma-separated file "ChickWeight.csv". 
```{r}
# Note: this dataset can also be accessed directly from the ChickWeight package in R 
# (see ?ChickWeight)
chick = read.table("../ChickWeight_Challenge_EASY/ChickWeight.csv", header = T, sep = ",")
```

### First, explore the data. \newline
How many chicks are in the dataset? \newline
How many different diets are in the experiment? \newline
```{r}
length(unique(chick$Chick))
length(unique(chick$Diet))
```

### What are the range of chick weights in the dataset?
```{r}
range(chick$weight)
```

### What chick was the heaviest? Lightest?
```{r}
chick$Chick[which.max(chick$weight)] # heaviest
chick$Chick[which.min(chick$weight)] # lightest
# Equivalently could do chick$Chick[which(chick$weight == max(chick$weight))]
```

### To vizualize the basics of the data, plot weight versus time
```{r}
plot(chick$weight ~ chick$Time, 
     xlab = "Time (days)", 
     ylab = "Weight (gm)")
```

### Create a multipanel plot of the first four chick weight versus time \newline
Hint: Use par()
```{r}
par(mfrow = c(2,2))
plot(chick$weight[chick$Chick == 1] ~ chick$Time[chick$Chick == 1], 
     type = "l", main = "Chick 1", xlab = "Time (days)", ylab = "Weight (gm)", col = "red")
plot(chick$weight[chick$Chick == 2] ~ chick$Time[chick$Chick == 2], 
     type = "l", main = "Chick 2", xlab = "Time (days)", ylab = "Weight (gm)", col = "blue")
plot(chick$weight[chick$Chick == 3] ~ chick$Time[chick$Chick == 3], 
     type = "l", main = "Chick 3", xlab = "Time (days)", ylab = "Weight (gm)", col = "purple")
plot(chick$weight[chick$Chick == 4] ~ chick$Time[chick$Chick == 4], 
     type = "l", main = "Chick 4", xlab = "Time (days)", ylab = "Weight (gm)", col = "orange")

# This can also be accomplished with a for loop:
for (i in 1:4){ # for each chick i
  plot(chick$weight[chick$Chick == i] ~ chick$Time[chick$Chick == i], 
       type = "l", 
       main = paste("Chick", i), # Chick 1, Chick 2 etc.
       xlab = "Time (days)", 
       ylab = "Weight (gm)",
       col = "darkblue")
}
```

### Plot a histogram of the weights of the chicks at the final day of the experiments (i.e. only the chicks who made it to the last day)
```{r}
par(mfrow = c(1,1))
hist(chick$weight[chick$Time == max(chick$Time)], 
     xlab = "Weight (gm)", 
     main = "Weights at final day of experiment", 
     col = rainbow(10))
```

### Create a boxplot where the x-axis represents the different diets and the y-axis is the weights of the chicks at the final day of the experiments
```{r}
my.new = chick[chick$Time == max(chick$Time), ]
boxplot(weight ~ Diet, 
        data = my.new, 
        xlab = "Diet", 
        ylab = "Weight (gm)", 
        main = "Final weights given diet type", 
        col = c("red", "blue", "orange", "green"))
```

### Find the last time entry for each chick in the dataset and plot a histogram of these times
```{r}
lastTime = sapply(unique(chick$Chick), function(i) 
  max(chick$Time[chick$Chick == i]))
hist(lastTime, 
     xlab =  "Last time entry (days)", 
     main = "Histogram of last time entries", 
     col = rainbow(length(unique(lastTime))))
```

### Create a function that adds points to a existing plot for an individual chick's weight over time
- As arguments for the function be just the individual chick to be plotted
- In plotting, change "type" to plot both lines and points
- Have the color argument be able to take a vector of colors which length diet type (that you will make later) so that each diet has the same color across chicks
Hint: You can update this function later - if you're having trouble with coloring by diet, try plotting everything the same color first
```{r}
plotChickWeight = function(individual) {
  points(chick$Time[chick$Chick == individual], 
         chick$weight[chick$Chick == individual], 
         type = "b", 
         col = colors[chick$Diet[chick$Chick == individual]])
}

# Now, create an empty plot to add these points to 
# (hint: can recreate the first plot with "col = NULL")
with(chick, plot(Time, weight, 
                 col = NULL, 
                 main = "Chick weight through time",
                 ylab = "Weight (gm)", 
                 xlab = "Time (days)"))

# Create a vector of color values to pass the function, 
# so that each diet type has a different color line/points
# Useful functions: rainbow()
colors = rainbow(length(unique(chick$Diet)))

# Now, use your function to add lines/points to the plot for all chicks
my.out = sapply(unique(chick$Chick), plotChickWeight)

# Overlay a solid, thick red line of average weight (across all chicks) 
# at each time interval on top of this graph
meanWeight = tapply(chick$weight, chick$Time, FUN = mean)
lines(x = unique(chick$Time), meanWeight, 
      type = "l", col  = "red", lwd = 3)

#Add a legend to specify what the solid, thick red line represents
legend("topleft", lty = 1, 
       col  = "red", lwd = 3, legend = "Mean weight")
```

### Try using the package R Color Brewer to generate color palettes. Go to http://colorbrewer2.org/ to vizualize palettes. You can choose palettes that are colorblind safe, print friendly, etc.
```{r}
# Install R Color Brewer
# install.packages("RColorBrewer")
library("RColorBrewer")
```
## Examples of R Color Brewer:
### Define color pallete with 10 colors and re-plot histogram of histogram of the weights of the chicks at the final day of the experiments in these colors \newline
Note: if histogram has n breaks and n is less than 10, it will just use first n colors. If n is greater than 10, it will reuse colors.
```{r}
my.colors = brewer.pal(10, "Paired")
hist(chick$weight[chick$Time == max(chick$Time)], 
     xlab = "Weight (gm)", 
     main = "Weights at final day of experiment", 
     col = my.colors)
```

### Re-do final plot with new colors and solid lines instead of points and lines
```{r}
plotChickWeight_newColors = function(individual) {
  lines(chick$Time[chick$Chick == individual], 
        chick$weight[chick$Chick == individual], 
        col = new.colors[chick$Diet[chick$Chick == individual]])
}

with(chick, 
     plot(Time, weight, 
          col = NULL, 
          main = "Chick weight through time",
          ylab = "Weight (gm)", 
          xlab = "Time (days)"))
new.colors = brewer.pal(length(unique(chick$Diet)), "Set2")

my.out = sapply(unique(chick$Chick), plotChickWeight_newColors)
```

### Now, let's put this into R Markdown, a package that makes nice final reports of your code and figures \newline
Go back and make sure your plots have interpretable axes labels and titles. \newline
If you're interested, change the colors and point/line types to make your figures more vizually appealing. \newline
```{r}
# Install R Markdown
# install.packages("rmarkdown")
library("rmarkdown")
```
### Use R Studio to create new R Markdown file with the relevant code and output that you generated above. Create an HTML or PDF document.
