---
title: "Earthquake Challenge"
output:
  pdf_document: default
---

## Should you be worried about earthquakes? Let's make a map.
#### useful functions (what I used to do this): dim(), distm(), install.packages(), library(), map(), mapply(), mapproject(), max(), min(), function(){}, points(), read.csv()

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Let's make a map. Make a map in R. Use data from https://earthquake.usgs.gov/earthquakes/search/ to generate a map of earthquakes.
#### Packages
```{r}
# installing all the packages I'll use
# install.packages(c("maps","mapproj","geosphere")) # uncomment and run once
library(maps)
library(mapproj)
library(geosphere)
```
#### Generate dataset
- get data from https://earthquake.usgs.gov/earthquakes/search/
- download a CSV from this site (you select the time period) I used the last 30 days

#### Import earthquake data in R
```{r}
recentquakes <- read.csv("~/Downloads/query.csv", stringsAsFactors=F)
# you'll need to put your own file here
```

#### Map all the recent earth quakes. 
```{r}
# get a map
par(mar=c(0,0,0,0))
map("world")

# get points on the map
ptsproj <- mapproject(recentquakes$longitude, recentquakes$latitude)
points(ptsproj)
```

#### How big was the biggest earthquake?
```{r}
max(recentquakes$mag)
```

#### How far was your home from the biggest earthquake? 
```{r}
# where do you live?
# my (approximate) address where I used to live
mylat = 30.35
mylong = -97.75

# orders earthquakes biggest to smallest
orderedquakes <- recentquakes[order(recentquakes$mag, decreasing = T), ]

# biggest earthquake
biggest <- orderedquakes[1, ]
biggest

# note: this can also be solved this way, 
# but we still use [1, ] to take just 1 row in case there are ties
# biggest quake:
recentquakes[recentquakes$mag == max(recentquakes$mag), ][1, ]

# how big was it?
biggest$mag

# distance from home (m)
distm(c(mylong, mylat), c(biggest$long, biggest$lat))
```

#### Write a function to calculate distance from your home
```{r}
fromhome <- function(lat, long) {distm(c(mylong, mylat), c(long, lat))}
```

### Add the distance from home as a column in the original dataframe. Then identify the closest recent earthquake. What was it's magnitude?
```{r}
recentquakes$distancesfromhome <- mapply(fromhome,
                                  recentquakes$latitude,
                                  recentquakes$longitude)
# Closest earthquake
closestquake <- recentquakes[recentquakes$distancesfromhome ==
                               min(recentquakes$distancesfromhome), ]

# Look at the magnitude
closestquake$mag
```


#### Improve your map
- Play around with colors to make your map appealing and readable. 
- Map all the recent earth quakes.
- Put the biggest earthquake on the map with a different symbol and color.
- Put your home on the map, too, in another different color and symbol.
- Then add the earthquake closest to your home, again in a different color.
```{r}
# get a map
par(mar=c(0,0,0,0))
map("world", 
    col = "blue", 
    bg = "#000050", 
    fill = FALSE, 
    interior = TRUE, 
    lwd = 0.5, 
    projection = "cylequalarea", 
    par = 0, 
    wrap = TRUE)

# get points on the map
ptsproj <- mapproject(recentquakes$longitude, recentquakes$latitude)
points(ptsproj, pch=20, cex=.5, col="white")

# Biggest earthquake on the map
biggest_loc <- mapproject(biggest$long, biggest$lat)
points(biggest_loc, cex=.8, pch=15, col="yellow")

# get home on the map
# put on the map with a star
myloc <- mapproject(mylong, mylat) # home longitude and latitude
points(myloc, cex=.8, pch=10, col="red")

# get the coordinates for the closest earthquake
closequake <- mapproject(closestquake$long, closestquake$lat)
# and plot it in a different color
points(closequake, cex=.8, pch=20, col="green")
```

#### How many earthquakes occured within a long days drive of you (500 miles)? How big was the biggest one of these?
```{r}
# find all the close earthquakes
# 500 miles= 800,000 m
quakes500close <- recentquakes[(recentquakes$distancesfromhome) < 800000, ]

dim(quakes500close) # number of earthquakes close to me

max(quakes500close$mag) # biggest earthquake
```