---
title: "Exercise 3.1"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## For loops and sapply()
#### useful commands: `for(){ }`, `print`, `paste`, `sample`, `plot`, `unique`, `points`, `rep`, `which`

#### 01*. Create a for loop to print "The year is xxxx" for 2010-2016.

```{r}
for (i in 2010:2016) { # for each year i
  # 'paste' the year into the sentence, then 'print' to console
  print(paste("The year is", i, sep = " ")) 
}
```


#### 02*. Instead of printing the answer from 01, save it as a character vector called 'my_years'. 
# You could do this with a for loop, but that would more difficult and less efficient, so try using sapply()
```{r}
my_years = sapply(2010:2016, FUN = function(i) paste("The year is", i, sep = " "))
```

#### 03*. Use a for loop to solve: Find the 100th number in the Fibonacci series: 1, 1, 2, 3, 5, 8, 13, 21 ... 
```{r}
fib = rep(NA, 100) # make a vector length 100 of NA's to store the results
fib[1] = 1 # set the first value of the vector = 1
fib[2] = 1 # set the second value of the vector = 1
# your for loop to fill in the rest of the fib vector
for (i in 3:100){ # starts at 3 because first 2 Fibonacci numbers are given
  # fills in the next number in the series using the Fibonacci pattern
  fib[i] = fib[i-1] + fib[i-2] 
}
# the answer
fib[100] # 100th Fibonacci number
# my self-check tests
fib[8] # 8th Fibonacci number
fib[c(1:15)] # first 15 Fibonacci numbers
```

#### BONUS 03b. How many steps into the Fibonacci series until you get the first number over 10,000?
```{r}
min(which(fib > 10000)) # index of minimum Fibonacci number over 10000
fib[21] # the number
fib[min(which(fib > 10000))] # putting it together
```

#### BONUS 04. use sapply() to plot a histogram of the data below 4 times, in 4 different colors. For extra style, title the plot by it's color, e.g. the red plot is titled "red"

```{r}
data_to_plot = c(1,3,4,5,6,3,3,4,5,1,1,1,1,1)
my_colors = c("deeppink", "red", "blue", "darkgreen")
par(mfrow = c(2,2)) # extra styling, plots in a 2x2 grid
sapply(my_colors, FUN = function(i) 
  hist(data_to_plot, main = i, col = i))
```


#### BONUS 05. How would you do question 2 above (print "The year is xxxx" for 2010-2016) without using sapply() or a for loop?
```{r}
paste("The year is", 2010:2016, sep = " ") # vector short-hand
```
