---
title: "Exercise 3.2"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Apply family
#### useful commands: `function(){ }`, `apply()`, `tapply()`, `hist()`, `dim()`, `prod()`, `sd()`

#### 00*. Load state data
```{r}
?state
data(state)
# this data is stored in a slightly different way than other datasets we've used so far
states = as.data.frame(state.x77) # run this line of code to avoid later confusion
```


#### 01*. What is the average population, income, and area of all 50 states? \newline
Do this in one short line of code

```{r}
apply(states, 2, mean)
colMeans(states) # alternative
```

#### 02*. What is the average area of the states from different regions of the country? \newline
Hint: use the object state.region in your environment

```{r}
tapply(states$Area, state.region, mean)
```

#### 03*. Plot a histogram for each variable in the states data (Population, Income, Illiteracy etc.)
```{r, results = 'hide'}
# 'hide' code chunk option only shows plots, no text output
dim(states)
par(mfrow = c(2,2)) # shows 2 rows and 2 columns of plots at once in a frame
apply(states, 2, hist)
```

#### 04*. Let's assume that we don't want to live in a state with high illiteracy, high murder, and many freezing days. \newline 
Also assume that each of these factors contribute equally to our opinion: (Illiteracy x Murder x Frost) = undesirable. \newline
Which 10 states should we avoid? \newline
Hint: use prod(); and maybe order()
```{r}
livability = apply(states[,c("Illiteracy", "Murder", "Frost")], # subset to variables of interest
                   1, prod) # apply product (i.e. multiplication) to each row
livability[order(livability, decreasing = T)][1:10] # top ten least livable states
```

#### BONUS 05. Standardize all the variables in the states dataset and save your answer to a new dataframe, states_standardized \newline
Hint: to standardize a variable, you subtract the mean and divide by the standard deviation (sd)

```{r}
states_standardized = apply(states, 2, function(x) (x-mean(x))/sd(x))
# original:
head(states)
# new:
head(states_standardized)
```


#### BONUS 06. Create a histogram again for each variable in the states data, but this time label each histogram with the variable names when you plot \newline
Hint: instead of using apply to iterate over the columns themselves, you can often iterate over the column names with sapply
```{r, results = 'hide'}
par(mfrow = c(2,2))
sapply(colnames(states), function(x) 
  hist(states[ , x], 
       main = x, xlab = x,
       col = "darkblue"))
```
