---
title: "Exercise 2.1"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Plotting
#### useful commands: `hist()`, `barplot()`, `boxplot()`, `plot()` (for scatter plot)
#### useful arguments within plot(): `main`, `xlab`, `ylab`, `col`, `pch`, `cex`


##### 00*. read in the "crabs.csv" dataset

```{r}
getwd() 
# here the working directory is STEMinist_R/Day_2/Day_2_Solutions/, 
# and crabs.csv is located in STEMinist_R/Day_2/Day_2_Exercises/ . 
# You can change the working directory using setwd()
crabs <- read.csv("../Day_2_Exercises/crabs.csv", 
                  header = T,
                  stringsAsFactors = F)
# note: this dataset is published as part of the MASS package
```


##### 00*. subset the data to only include the blue crabs
```{r}
blueCrabs = crabs[crabs[ ,"sp"] == "B", ]

```


##### 01*. plot carapace length versus carapace width of the blue crabs

- 01a*. make a scatterplot using the 'plot' command
```{r}
plot(x = blueCrabs[ , "CW"], blueCrabs[ , "CL"])

```

- 01b*. change the color of the points
```{r}
plot(x = blueCrabs[ , "CW"], blueCrabs[ , "CL"], col = "red")

```
- 01c.change the shape of the points to a triangle
```{r}
plot(x = blueCrabs[ , "CW"], blueCrabs[ , "CL"], col = "red", pch = 2)

```

- 01d. change the size of the points
```{r}
plot(x = blueCrabs[ , "CW"], blueCrabs[ , "CL"], col = "red", pch = 2, cex = 2)
```


- 01e*. add a title and axis labels to the plot
```{r}
plot(x = blueCrabs[ , "CW"], blueCrabs[ , "CL"], col = "red", pch = 2, cex = 2, xlab = "Carapace width (mm)", ylab = "Carapace length (mm)", main = "Carapace length versus carapace width of blue crabs")

```

##### 02*. make a barplot of the number of observations for each sex of blue crabs
```{r}
barplot(table(blueCrabs[ , "sex"]))

```

##### 03*. make a histogram of the body depth of the blue crabs
```{r}
hist(blueCrabs[, "BD"])

```

##### 04*. boxplot of body depth by sex of the blue crabs
```{r}
boxplot(blueCrabs[ ,"BD"] ~ blueCrabs[ , "sex"])

```

- 04a* plot this in different colors for each sex
```{r}
boxplot(blueCrabs[ ,"BD"] ~ blueCrabs[ , "sex"], col = c("red", "blue"))

```

- 04b add a main title and axis labels
```{r}
boxplot(blueCrabs[ ,"BD"] ~ blueCrabs[ , "sex"], col = c("red", "blue"), main = "Blue crab body depth by sex", ylab = "Sex", xlab = "Body depth (mm)")

```


##### 05 BONUS** - defining color variables for plotting
 useful commands: [logical,]= , `ifelse()`, `plot()` + `points()`

make scatter plot for carapace length versus carapace width of the blue crabs

but now depict records of female crabs with red, and male crabs with blue

- Solution 1 --- add a column of colors to the original dataset using indexing
```{r}
# add a new column to the data.frame, and assign "dummy" values for now
blueCrabs$color = NA 
# now assign "red" to the new column when crabs are females
blueCrabs$color[blueCrabs$sex == "F"] = "red" 
# now assign "blue" when they are males
blueCrabs$color[blueCrabs$se == "M"] = "blue" 
# depict females and males with different colors
plot(blueCrabs$CL ~ blueCrabs$CW, col = blueCrabs$color)
```


- Solution 2 -- add a color column to the dataset using an "ifelse" statement
```{r}
blueCrabs$color2 = ifelse(blueCrabs$sex == "F", "red", "blue")
# if crabs are females add "red", but if they are not females, add "blue" to the new column

# plot mass across time, and depict females and males with different colors
plot(blueCrabs$CL ~ blueCrabs$CW, col = blueCrabs$color2)

```

- Solution 3 --- make NULL plot and add points  (note: this can be a good strategy to combine with sapply for a large number of categories)
```{r}
plot(blueCrabs$CL ~ blueCrabs$CW, col = NULL)
points(blueCrabs[blueCrabs$sex == "F", "CW"], blueCrabs[blueCrabs$sex == "F", "CL"], col = "red")
points(blueCrabs[blueCrabs$sex == "M", "CW"], blueCrabs[blueCrabs$sex == "M", "CL"], col = "blue")

```


