---
title: "Exercise 2.2"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### User generated functions
#### useful commands: `function(){}`, `is.na()`, `which()`, `length()`, `floor()`, `median()`, `%%` (remainder), `month.name`

#### To make a function, we use:\newline
my_function = function ( arglist )  {body}

##### 01*. create a function that turns 0's in a numeric vector into NA's
```{r}
zero_to_NA = function(v){
  v_new = v
  v_new[v==0] = NA
  return(v_new)
}

```

##### test your function by creating a vector that contains zeros and calling the function
```{r}
test1 = c(1,0,0,3)
zero_to_NA(test1)
```


##### 02*. create a function that plots a histogram of a vector and also adds a vertical line at the median

##### test your function by creating a numeric vector and calling the function

##### BONUS: use ... to pass optional arguments to your function (e.g. a title or color for your histogram)

```{r}
medHist = function(v, ...){ #the name and arguments specified
  #additionally, we added ... to accept any other graphical arguments
  hist(v, ...) # we pass the optional graphical parameters to hist() using ...
  abline(v = median(v), col = "red")
}

medHist(c(1,1,1,1,2,3,4,5), col = "blue", main = "my histogram") # testing function

```

##### BONUS 03. create a function that when given a month name
    - prints the subsequent month's names 
      - i.e. `nextMonth("January")` returns `"February"`
    - hint: the vector of month names is built into R .. try typing `month.name` in your console
    - make sure your function works for December!

```{r}
nextMonth = function(m){
  index = (which(month.name == m) + 1)%%12 
  # %% gets the remainder from dividing by 12 .. so '13' becomes 1 (January)
  new_month = month.name[index]
  return(new_month)
}

```

##### Can also use if statements
```{r}
nextMonth = function(m){
  index = (which(month.name == m) + 1)
  if(index == 13) {
    return(month.name[1])
  }
  if(index != 13) {
    new_month = month.name[index]
  }
}

nextMonth("February") # testing
nextMonth("December") # testing -> round corner back to January

```

##### note: can use `sapply()` to go over vector of more than length 1 (i.e. multiple months)

```{r}
sapply(c("February", "December"), nextMonth) #returns "March", "January"
```


